A molecular dynamics study of the solvation of CO\textsubscript{2}, water, and some organic compounds in the ionic liquids [emim][B(CN)\textsubscript{4}] and [emim][NTf\textsubscript{2}]

### 2018-04-04 ionic_liquids.tex
discusionI

In this paper, we consider four models of [emim][B(CN)\textsubscript{4}]. Koller~\textit{et al.}~\cite{Koller_2012} were the first to introduce a specific non-polarizable force field of [emim][B(CN)\textsubscript{4}]. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups $\text{CH}_3$ y $\text{CH}_2$ of [emim]$^{+}$ are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and Lennard-Jones parameters of the cation were taken from the work of Liu~\textit{et al.}~\cite{Liu_2006}, which in turn is based on the AMBER~\cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING~\cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price~\textit{et al.}~\cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method~\cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6–31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426e.

@@

With the aim of setting up a (more) computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method~\cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes, McCann~\textit{et al.}~\cite{McCann_2013} tested the DSF method for various ILs modeled in the OPLS-AA framework. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provides the best energetic agreement with Ewald. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in good agreement for the ILs studied here. Those values were also employed for the charged solutes, with the exception of hexane, for which we considered undamped Coulombic interactions.

@@

In this work, we performed independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the states associated to the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of the EE involves both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, the remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT MD for a determined number of steps $N^{\text{EE}}$ followed with an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. Independent sampling of $\lambda$~\cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$~\cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were all implemented in the LAMMPS software package~\cite{Plimpton1995}.

@@

We end this section with a brief description of the procedure for generating the initial configurations. Only Ref.~\citenum{Liu_2014} provides the atomic coordinates of the cation-anion pair corresponding to a minimum energy configuration obtained in an \textit{ab-initio} simulation. For the other models, we employed the tool \textit{build} of PLAYMOL~\cite{playmol}, a software integrated with PACKMOL~\cite{Mart_nez_2009}, which enables to specify the following geometric data: bond and angle lengths and dihedral angles, from which PLAYMOL automatically determines the Cartesian coordinates. We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation for the corresponding (classical) model of each pair. These simulations were performed using the the Nos\'{e}-Hoover thermostat~\cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions. To relax the initial configurations, a short run with a small time step of 0.1 fs was performed, which was followed by an equilibration using $\Delta t$ = 0.5 fs. Finally, we carried out a ``quenching'' simulation with $\Delta t$ = 1 fs, in which the temperature is gradually reduced until reaching a value close to 0 K. The resulting coordinates were specified in PLAYMOL to build the initial configurations for simulating the liquid phases.

@@

\subsection{Simulation strategies of ionic liquids}
\label{sec:prel_results}
In this section, we present results of preliminary simulations using the  alternatives to standard approaches of modeling the ions as rigid bodies and employing the DSF method~\cite{Fennell2006} For that, we compare the simulated properties of density and radial distribution functions to that obtained when considering the ions as flexible molecules and the Ewald method for the Coulombic interactions.

For the determining the initial configuration to specify for the rigid bodies we performed a preliminary study of the dihedrals in the ethyl group of [emim]$^{+}$, as we consider that its flexibility  may introduce a more significant conformational change in the cation. Small anions are usually simulated as rigid bodies. The nomenclature of the atoms are shown in Fig. In the \textit{united-atoms} version of [emim]$^{+}$,  the nomenclature is identical, but hydrogens ranging from H4 to H11 are eliminated.

Using the software TRAVIS~\cite{Brehm_2011}, for each model we obtain the combined distribution function for the angles of the dihedrals C1-N2-C5-C6 and C2-N2-C5-C6 sampled in NPT simulations of 250 ion pairs. We chose this analysis, in place of individual histograms, as it provides simultaneously rigidity of the ring. In Figure we present the results for each model, together withe the color scale associated to the histograms.

A first inspection of Fig.~\ref{fig:atoms_id} reveals a similar behavior of the models developed by Koller~\textit{et al.}~\cite{Koller_2012} and Liu~\textit{et al.}~\cite{Liu_2014}. This was expected as in both cases are based on AMBER. In addition, Figs are symmetric, which means that atoms C1-N2-C2 are coplanar.

@@

  \includegraphics[width=.4\linewidth]{Liu.png}%
  \includegraphics[width=.4\linewidth]{Weber.png}%
\caption{}
\label{k}
\end{figure*}

>>>>

In this paper, we consider four models of [emim][B(CN)\textsubscript{4}]. Koller~\textit{et al.}~\cite{Koller_2012} were the first to introduce a specific non-polarizable force field of [emim][B(CN)\textsubscript{4}]. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups $\text{CH}_3$ and $\text{CH}_2$ of [emim]$^{+}$ are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and Lennard-Jones parameters of the cation were taken from the work of Liu~\textit{et al.}~\cite{Liu_2006}, which in turn is based on the AMBER~\cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING~\cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price~\textit{et al.}~\cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method~\cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6–31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426e.

@@

With the aim of setting up a (more) computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method~\cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes~\cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann~\textit{et al.}~\cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provides the best energetic agreement with Ewald. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in good agreement for the ILs studied here. Those values were also employed for the charged solutes, with the exception of hexane, for which we considered undamped Coulombic interactions.

rigid treatment of the imidazolium ring~\cite{Kobrak_2004,Lynden_Bell_2006}
\cite{Rey_Castro_2007}

@@

In this work, we performed independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the states associated to the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of the EE involves both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, the remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT MD for a determined number of steps $N^{\text{EE}}$ followed with an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. Independent sampling of $\lambda$~\cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$~\cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were all implemented in the LAMMPS software package~\cite{Plimpton1995}, which we employed for carrying out all the MD simulations.

@@

We end this section with a brief description of the procedure for generating the initial configurations. Only Ref.~\citenum{Liu_2014} provides the atomic coordinates of the cation-anion pair corresponding to a minimum energy configuration obtained in an \textit{ab-initio} simulation. For the other models, we employed the tool \textit{build} of PLAYMOL~\cite{playmol}, a software integrated with PACKMOL~\cite{Mart_nez_2009}, which enables to specify the following geometric data: bond and angle lengths and dihedral angles, from which PLAYMOL~\cite{playmol} automatically determines the Cartesian coordinates. We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation for the corresponding (classical) model of each pair. These simulations were performed using the the Nos\'{e}-Hoover thermostat~\cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions. To relax the initial configurations, a short run with a small time step of 0.1 fs was performed, which was followed by an equilibration using $\Delta t$ = 0.5 fs. Finally, we carried out a ``quenching'' simulation with $\Delta t$ = 1 fs, in which the temperature is gradually reduced until reaching a value close to 0 K. The resulting coordinates were then specified in PLAYMOL~\cite{playmol} to build the initial configurations for simulating the liquid phases.

@@

\subsection{Simulation strategies of ILs}
\label{sec:prel_results}

In this section, we present and compare the results obtained for the following two cases: (1) flexible ions and the Particle-Particle-Particle-Mesh (PPPM) method~\cite{Hockney_1988} with a tolerance of 1$\times$10$^{-4}$ for the evaluation of the Ewald sums~\cite{Ewald_1921}, and (2) ions modeled as rigid bodies with the DSF method~\cite{Fennell2006} for calculating the Coulombic interactions.

First, we describe the approach for defining the rigid bodies. For that, we carried out an analysis of key dihedrals, particularly those whose flexibility could lead to significant conformational changes in [emim]$^{+}$ and [Tf\textsubscript{2}N]$^{-}$. On the other hand, small and highly symmetric anions, such as PF\textsubscript{6} and BF\textsubscript{4} are usually modeled as rigid bodies~\cite{HANKE_2001,Hanke_2003,Lynden_Bell_2006}, which justifies that strategy in the case of [B(CN)\textsubscript{4}]$^{-}$. The atomic nomenclatureare adopted in this paper is shown in Fig~\ref{fig:atoms_id}. In the \textit{united-atoms} version of [emim]$^{+}$, the nomenclature is identical, but the hydrogens ranging from H4 to H11 are eliminated.

Using the software TRAVIS~\cite{Brehm_2011}, for each model we obtain the combined distribution function (CDF) of the dihedral angles containing the ehtyl chain atoms, which are C1-N2-C5-C6 and C2-N2-C5-C6, as sampled in NPT simulations of 250 ion pairs. The CDF, as an alternative of individual histograms, provides a way to  simultaneously analyse the rigidity of the ring. In Fig.~\ref{fig:dihedrals} we present the results of CDF for each model, together with the color scale associated to the histograms.

A first inspection of Fig.~\ref{fig:atoms_id} reveals a similar behavior of the models developed by Koller~\textit{et al.}~\cite{Koller_2012} and Liu~\textit{et al.}~\cite{Liu_2014}. This was expected as both were developed based on AMBER force field. In addition, Figs are symmetric, which means that atoms C1-N2-C2 are coplanar.

@@

  \includegraphics[width=.4\linewidth]{Liu.png}%
  \includegraphics[width=.4\linewidth]{Weber.png}%
\caption{}
\label{fig:dihedrals}
\end{figure*}
